PhD Title \\ PhD Thesis

### 2024-07-04 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

Vaterite samples where made by the first preparing equal amounts of 
$CaCl_2$ and $NaCO_3$ at a concentration of $0.1M$, at the same time
a vial of $0.03M\ MgSO_4$ was prepared and set aside for later. First
a small vial was filled with $1.5mL$ of $CaCl_2$ followed by $60\mu L$ 
and $90\mu L$ of $MgSO_4$ and $NaCO_3$ respectively, forming a seed solution. 
Next, a larger vial was filled with $5\ mL$, $1.5\ mL$, and $1\ mL$ of
$CaCL_2$, $MgSO_4$, and $NaCO_3$ respectively followed by the seed solution. 
After 10 minutes of slow but continuous mixing a few drops of Agepon was added
to halt the reaction, the solution was filtered and washed 3 times with 
distilled water before being suspended in water.

\subsection{Rotation of Vaterite micro spheres}
Vaterite is a polymorph of calcium carbonate that is rarely seen in 
nature due to its low stability. However unlike its other polymorphs of 
calcite and aragonite, when synthesised vaterite will typically form 
small spherical particles making them ideal for optical trapping. 
Synthesis of vaterite micro spheres requires fine control of the nucleation process in order to maintain polymorphic control.

>>>>

Vaterite is a polymorph of calcium carbonate that is rarely seen in 
nature due to its low stability. However unlike its other polymorphs of 
calcite and aragonite, when synthesised vaterite will typically form 
small spherical particles making them ideal for optical trapping. 
Synthesis of vaterite micro spheres requires fine control of the nucleation
process in order to maintain polymorphic control.

Vaterite samples where made by the first preparing equal amounts of 
$CaCl_2$ and $Na_2CO_3$ at a concentration of $0.33M$, at the same time
a vial of $0.33M\ MgSO_4$ was prepared and set aside for later. First
a small vial was filled with $1.5mL$ of $CaCl_2$ followed by $60\mu L$ 
and $90\mu L$ of $MgSO_4$ and $NaCO_3$ respectively, forming a seed solution. 
Next, a larger vial was filled with $5\ mL$, $1.5\ mL$, and $1\ mL$ of
$CaCL_2$, $MgSO_4$, and $NaCO_3$ respectively followed by the seed solution. 
After 10 minutes of slow but continuous mixing a few drops of Agepon was added
to halt the reaction, the solution was filtered and washed 3 times with 
distilled water before being suspended in water.

\begin{figure}
	\centering
	\begin{subfigure}{0.4\linewidth}
		\includegraphics[width=\linewidth]{vaterite_sample.jpg}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.55\linewidth}
		\includegraphics[width=\linewidth]{rotating_psd.png}
		\subcaption{}
	\end{subfigure}
	\caption{(a)Sample vaterite sphere suspended in water and trapped by circular polarised trap. (b) Collected power spectrum from rotating vaterite, peaks in the power spectrum appear at integer multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
\end{figure}

\subsection{Rotation of Vaterite micro spheres}

### 2024-07-04 Latex_files/Thesis/chapters/chapter 4/main.tex
updates

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions. (a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position. (b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
\end{figure}
\newpage
\subsection{Non-trivial harmonic traps}\label{sec:off-axis}
Computing the equilibrium positions when a dimer is aligned with the electric field is relatively simple as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap. However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}. Unlike before where we can simply find the trap by varying the dimer's vertical position its instead more prudent to run a multitude of smaller simulations at a variety of starting positions and orientations. An example for a dimer of size ratio 2 is shown below:
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.67\linewidth}
		\includegraphics[width=\linewidth]{off_axis_trap.png}
	\end{subfigure}
	\begin{subfigure}{0.32\linewidth}
		\raisebox{-110pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}

		\caption{}
	\end{subfigure}
	\captionsetup{hangindent=2.35cm}
	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW. The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in. Right hand render shows a dimer in its off-axis configuration}
\end{figure}

@@

As shown in fig.~\ref{fig:psd_sphere}, the two power spectra
report different corner frequencies which would be indicate 
that the trap is not perfectly circular. Using \textit{ott} 
we can compare the expected trap strength to what is reported
by a quadrant photo diode:
\begin{center}
	\begin{tabular}{ |c|c|c|c|c| } 
		\hline
		Trap parameter & \textit{ott} x-axis & \textit{ott} y-axis & QPD x-axis & QPD y-axis \\
		\hline
		$f_c\ [Hz]$ & 447 & 450 & 439 & 474 \\
		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 \\
		\hline
		Ellipticity &
		\multicolumn{2}{|c|}{8.16 \%} &
		\multicolumn{2}{|c|}{27.17 \%} \\
		\hline
	\end{tabular}
\end{center}

@@

\begin{center}
	\begin{tabular}{ |c|c|c|c|c| } 
		\hline
		Trap parameter & \textit{ott} x-axis & \textit{ott} y-axis & QPD x-axis & QPD y-axis \\
		\hline
		$f_c\ [Hz]$ & 409 & 334 & 431 & 424 \\
		$\kappa\ [pN/\mu m]$ & 48.51 & 39.58 & 51.13 & 50.26 \\
		\hline
		Ellipticity &
		\multicolumn{2}{|c|}{42.8 \%} &
		\multicolumn{2}{|c|}{12.7 \%} \\
		\hline
	\end{tabular}
\end{center}

>>>>

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions. (a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position. (b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
\end{figure}
\newpage
\subsection{Non-trivial harmonic traps}\label{sec:off-axis}
Computing the equilibrium positions when a dimer is aligned with the electric field is relatively simple as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap. However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}. Unlike before where we can simply find the trap by varying the dimer's vertical position its instead more prudent to run a multitude of smaller simulations at a variety of starting positions and orientations. An example for a dimer of size ratio 2 is shown below:
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.67\linewidth}
		\includegraphics[width=\linewidth]{off_axis_trap.png}
	\end{subfigure}
	\begin{subfigure}{0.32\linewidth}
		\raisebox{-110pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
		\caption{}
	\end{subfigure}
	\captionsetup{hangindent=2.35cm}
	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW. The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in. Right hand render shows a dimer in its off-axis configuration}
\end{figure}

@@

As shown in fig.~\ref{fig:psd_sphere}, the two power spectra
report different corner frequencies which would be indicate 
that the trap is not perfectly circular. Using \textit{ott} 
we can compare the expected trap strength to what is reported
by a quadrant photo diode:

\begin{center}
	\begin{tabular}{ |c|c|c|c|c|c|c| } 
		\hline
		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} & \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Simulation fitting}\\
		\hline
		$f_c\ [Hz]$ & 447 & 450 & \parbox{1cm}{\centering 439} & 474 
		& \parbox{1.25cm}{\centering 523} & 513 \\
		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 & 61.94 & 60.7 \\
		\hline
		Ellipticity &
		\multicolumn{2}{|c|}{8.16 \%} &
		\multicolumn{2}{|c|}{27.17 \%} &
		\multicolumn{2}{|c|}{13.8 \%} \\
		\hline
	\end{tabular}
\end{center}

@@

\begin{center}
	\begin{tabular}{ |c|c|c|c|c|c|c| } 
		\hline
		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimate} & \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Simulation fitting} \\
		\hline
		$f_c\ [Hz]$ & 409 & 334 & \parbox{1cm}{\centering 431} & 424 
		& \parbox{1.25cm}{\centering 274} & 285 \\
		$\kappa\ [pN/\mu m]$ & 48.51 & 39.58 & 51.13 & 50.26 & 32.45 & 33.75 \\
		\hline
		Ellipticity &
		\multicolumn{2}{|c|}{42.8 \%} &
		\multicolumn{2}{|c|}{12.7 \%} & 
		\multicolumn{2}{|c|}{13.8 \%} \\
		\hline
	\end{tabular}
\end{center}
